\documentclass[12pt, twoside, a4paper]{report}

\usepackage[sort&compress]{natbib}
\usepackage[toc, page]{appendix}
\usepackage[nottoc]{tocbibind}

\begin{document}

\title{Real-time Pedestrian Modelling: Implementing the Ensemble Kalman Filter
for an Agent-Based Model}
\author{Keiran Suchak --- 200888140}
\maketitle

%\begin{abstract}
%This is the abstract.
%\end{abstract}

\tableofcontents
\listoffigures
\listoftables

\chapter*{Notes}

\begin{itemize}
    \item What is data assimilation?
    \item What is the Ensemble Kalman Filter?
    \item What are Agent-Based Models?
    \item How do people model pedestrians?
    \item EnKF: options for generating observations:
        \begin{itemize}
            \item External (the old way):
                \begin{itemize}
                    \item observations come from a previous run of the model
                    \item therefore, they are independent
                    \item the problem here is that the ensemble members probably
                        have been set with the wrong entrances and exits for
                        agents
                    \item this is likely more realistic, ie when modelling
                        people we ultimately won't know where they intend to go
                    \item but I don't think that we can overcome this without
                        parameter estimation DA.
                \end{itemize}
            \item Internal (the new way):
                \begin{itemize}
                    \item observations come from the \texttt{base\_model}
                    \item all ensemble members are deep copies of the
                        \texttt{base\_model}
                    \item consequently, they have the same parameters and
                        initial conditions
                    \item the problem with this is that is not realistic
                    \item but we're going to do it anyway (for now)
                \end{itemize}
        \end{itemize}
    \item What happens when an agent in one of the ensemble member models tries
        to leave the station early? Because we're only doing state estimation
        (meaning just the agent coordinates), the agent won't get reactivated.
    \item Errors should converge on something relating to twice the exit
        distance.
\end{itemize}

\section*{Structure}

\begin{itemize}
    \item What is the problem?
    \item Why should people care?
    \item What has been done in the past to try to solve the problem?
    \item What are we doing to solve the problem? How is this different?
    \item How well does our attempt work? How does this compare to others?
\end{itemize}

\include{chapters/introduction}

\include{chapters/lit_review}

\include{chapters/method}

\include{chapters/results}

\include{chapters/conclusion}

\begin{appendix}
    \include{appendices/code_documentation}
\end{appendix}

\bibliographystyle{agsm}
\bibliography{references}


\end{document}
