\documentclass[11pt,a4paper]{article}

\usepackage[square,sort,comma,numbers]{natbib}

\begin{document}

\title{Literature Review}
\author{Keiran Suchak}

\maketitle
\tableofcontents

\newpage

What is this section about?

\begin{itemize}
    \item Pedestrian modelling, and the use of agent-based modelling to achieve
        this
    \item Methods to analyse agent-based models and ensure that they are
        accurate
    \item The work that currently exists regarding the application of data
        assimilation methods to agent-based models, and the gap(s) in this field
        that we aim to fill
\end{itemize}

\section{Pedestrian Modelling}
\label{sec:pedestrians}

\begin{itemize}
    \item What is pedestrian modelling?
    \item Why should we care?
\end{itemize}

\section{Agent-Based Modelling}
\label{sec:abms}

\begin{itemize}

    \item How does agent-based modelling fit into this?
    \item What is agent-based modelling?
    \item Why should we use it?
\end{itemize}

\subsection{Model Description}
\label{sub:abms:description}

\begin{itemize}
    \item How do we go about describing ABMs?
        \begin{itemize}
            \item Systematically in words (ODD) \citep{grimm2006standard}
            \item Computationally/algorithmically \citep{north2014theoretical}
            \item Mathematically
                \citep{hinkelmann2011mathematical,hinkelmann2011adam}
        \end{itemize}
\end{itemize}

\subsection{Model Design Concepts}
\label{sub:abms:design}

\begin{itemize}
    \item What concepts feed into our model design process?
        \citep{railsback2011agent}
        \begin{itemize}
            \item Emergence
            \item Observation
            \item Sensing
            \item Adaptive behaviour
            \item Prediction
            \item Interaction
            \item Scheduling
            \item Stochasticity
            \item Collectives
        \end{itemize}
\end{itemize}

\subsection{Limitations}
\label{sub:abms:limitations}

\section{Model Analysis}
\label{sec:analysis}

\subsection{Validation}
\label{sub:analysis:validation}

\begin{itemize}
    \item \citep{xiang2005verification}
    \item \citep{klugl2008validation}
    \item \citep{ormerod2009validation}
\end{itemize}

\subsection{Sensitivity Analysis}
\label{sub:analysis:sensitivity}

\begin{itemize}
    \item What does it mean to perform sensitivity analysis?
    \item Why do we do it?
    \item How do we do it? \citep{thiele2014facilitating,ten2016sensitivity}
\end{itemize}

\subsection{Calibration}
\label{sub:analysis:calibration}

\begin{itemize}
    \item Calibration of ABMs
        \begin{itemize}
            \item Different methods of model calibration
                \citep{thiele2014facilitating}
        \end{itemize}
    \item Pros and cons of different methods
\end{itemize}

\section{Data Assimilation for Agent-Based Models}
\label{sec:dyn_cal}

\begin{itemize}
    \item What work has been done so far with data assimilation for ABMs
        \citep{ward2016dynamic,wang2015data}
        \begin{itemize}
            \item How valid/useful is this work?
            \item What issues are there with this work?
        \end{itemize}
    \item What gaps exist in the field?
    \item What can we do about them?
\end{itemize}

\bibliographystyle{unsrt}
\bibliography{ref}
\label{bib}

\end{document}
